---
title: "Assignment_2"
author: "Nicolas Crespo"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment 2
1. Add 2 + 2

```{r addition}
2 + 2
```

2. Create a list `dat` that goes from 1 to 10
3. View `dat`
4. Find the mean of `dat`

```{r dat}
# dat <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10) #this also works # nolint
(dat <- seq(1, 10, 1))
mean(dat)
```

5. Create another list called `dat2` that takes the sine of each 
individual element of dat multiplied by $\frac{\pi}{2}$.
```{r dat2}
dat2 <- sin(dat * (pi / 4))
dat2
```


6. View `dat2`
7. Plot `dat2` vs `dat` in a scatterplot
8. Plot `dat2` vs `dat` in a line graph
```{r Plot dat2 vs. dat}
plot(
  x = dat, y = dat2,
  main = "Title of scatterplot",
  xlab = "X-axis title",
  ylab = "Y-axis title"
)
plot(dat, dat2, type = "l")
```

9. Create a list called `dat3` with 5 A’s and 5 B’s
10. View `dat3`
```{r}
dat3 <- c("A", "A", "A")
```

11. Find the mean of `dat3`.
```{r dat3}
(dat3 <- c(rep("A", 5), rep("B", 5)))
mean(dat3)
```
```{r Plot dat vs. dat, eval = FALSE}
plot(dat3, dat, EVAL = FALSE)
```

13. Using the aggregate function, find the average of dat based on the list of A’s and B’s.
14. Make the code in [13] assigned to a variable called aggDat
15. View aggDat
```{r Aggregate function}
aggregate(dat, list(dat3), mean)
aggDat <- aggregate(dat, list(dat3), mean)
aggDat
```
16. Select the column x in aggDat
```{r Column select aggDat}
aggDat$x
```
17. Make a barplot of aggDat column x with “A” and “B” as the titles of each of the bars.
```{r Barplot of aggDat}
barplot(aggDat$x, names.arg = c("A", "B"))
```
